Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-09-28 paper_comparison/paper_comparison/paper_comparison.Rmd
modify exp design

As pointed out by @vanderplas2021statistical, the binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. They summarized three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design. For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$. Thus, the visual p-value followed by the above model is given as

@@

Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor. The same model was applied to different subsets of the survey data to make comparison of the power.

@@

## Motivation and overview

With our knowledge, the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics has not been validated. Particularly, its ability to detect non-linearity and heteroskedasticity compared to model structural test and Breusch–Pagan test. Meanwhile, as mentioned in Section \ref{introduction} section, we keen to understand the differences between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics and why conventional tests are not preferred by data analysts in linear regression diagnostics. Attempting to find the answer to these questions, three experiments were conducted.

The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test. The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test. The experiment III is designed for collecting human responses to lineup with only good residual plots such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated.

Subjects for all three experiments were recruited from an crowdsourcing platform called Prolific [ref here]. Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions.

During the experiment, every subject was presented with a block of 20 lineups. For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. An additional horizontal red line was added at $y = 0$ as a helping line. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others. If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason. No subject was shown the same lineup twice. Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.

In every block of 20 lineups that presented to a subject, two lineups with obvious visual patterns were included as attention checks. A subject's submission was only accepted if the actual data plot was identified for at least one attention check. Data of rejected submissions were discarded automatically to maintain the overall data quality.

Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects.

@@

where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as

@@

Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will result in model misspecification.

```{=tex}
\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}
```
where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$ and $g(.)$ is the scaling function defined in \ref{eq:scaling-function}.

For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $x$, which will lead to the presence of heteroskedasticity.

> Experiment I

-   Treatments for nonlinearity experiment (plots summarising levels)

Data of experiment I were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300), four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.

```{r results='asis'}
data.frame(n = c(50, 100, 300, ""),
           j = c(2, 3, 6, 18),
           sigma = c(0.5, 1, 2, 4),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
  knitr::kable(col.names = NULL,
               align = "rrrl",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$', 
               label = "parameter-table" ) %>%
  kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym2", "$\\\\sigma$", .) %>%
  gsub("sym3", "$n$", .) %>%
  gsub("sym4", "$X_{raw}$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

@@

The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. The example residual plots are given in Figure \ref{fig:different-sigma}.

```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

@@

Four different distribution were used to generate $X_{raw}$ shown in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density. And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I. This lineup was produced under $n = 300$, $j = 6$, $\sigma = 1$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup.

@@

@@

> Experiment II

-   Treatments for heteroskedasticity experiment (plots summarising levels)

Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.

The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include triangle shape, butterfly shape and inverse triangle shape.

```{r different-shape-of-heter}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("Triangle shape")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("Inverse triangle shape")) +
  theme_light() -> p3

The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.

@@

Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups. For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. Figure \ref{fig:example-lineup} is an example of one of these lineups. It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup.

In addition, each lineup is designed to be evaluated by five different subjects to provide reasonable estimates of the visual p-value. Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the design of the experiment I.

@@

>>>>

As pointed out by @vanderplas2021statistical, the binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. They summarized three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design. For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$. Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as

@@

(1) Instead, we modelled the probability of a lineup being rejected directly using a logistic regression with the natural logarithm of the effect size as the only regressor. The same model was applied to different subsets of the survey data for making comparison of the power.

(2) Instead, we modelled the plot selection $c_i$ for the actual data plot $i$.

@@

## Overview

<!-- With our knowledge, the effectiveness of visual test relative to the equivalent conventional test in regression diagnostics has not been validated. Particularly, its ability to detect non-linearity and heteroskedasticity compared to model structural test and Breusch–Pagan test. Meanwhile, as mentioned in Section \ref{introduction} section, we keen to understand the differences between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics and why conventional tests are not preferred by data analysts in linear regression diagnostics. Attempting to find the answer to these questions, three experiments were conducted. -->

Three experiments were conducted to investigate the difference between conventional hypothesis testing and graphical diagnostics in the application of linear regression diagnostics. The experiment I has ideal scenario for conventional testing, where the visual test is not expected to outperform the conventional test. The experiment II is a scenario where the conventional test is an approximate test, in which the visual test may have a chance to match the performance of the conventional test. The experiment III is designed for collecting human responses to lineup with only good residual plots such that the parameter $\alpha$ in Equation \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups performed by 443 subjects throughout three experiment.

@@

where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as

@@

Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will result in model misspecification. Visual patterns of non-linearity were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.

<!-- A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}. -->

```{r results='asis', eval = FALSE}
data.frame(j = c(2, 3, 6, 18),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
  knitr::kable(col.names = NULL,
               align = "rl",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Parameter values for $j$ and $X_{raw}$', 
               label = "parameter-table" ) %>%
  kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", "Distribution of sym4\n")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym4", "$X_{raw}$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

@@

<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. The example residual plots are given in Figure \ref{fig:different-sigma}. -->

```{r different-sigma, eval = FALSE, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

@@

Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density. And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I. This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup.

@@

```{=tex}
\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}
```

where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vectors of size $n$ and $g(.)$ is the scaling function defined in \ref{eq:scaling-function}.

For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $x$, which will lead to the presence of heteroskedasticity. Visual patterns were simulated using three different shapes ($a$ = -1, 0, 1) and the same four different distribution of $X_{raw}$ used in experiment I.

The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include left-triangle shape, butterfly shape and right-triangle shape as displayed in Figure \ref{fig:different-shape-of-heter}.

```{r different-shape-of-heter, fig.cap="Example residual plots of the linear regression model used in experiment II. Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle shape, butterfly shape and right-triangle shape."}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("Left-triangle shape")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("Right-triangle shape")) +
  theme_light() -> p3

To control the strength of the signal, two parameters were used including $\sigma$ and $b$ for controlling the standard deviation of the error term in experiment I and II respectively.

<!-- > Experiment I -->

<!-- -   Treatments for nonlinearity experiment (plots summarising levels) -->

@@

<!-- > Experiment II -->

<!-- -   Treatments for heteroskedasticity experiment (plots summarising levels) -->

@@

@@

Subjects for all three experiments were recruited from an crowdsourcing platform called Prolific [ref here]. Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions. During the experiment, every subject was presented with a block of 20 lineups. For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. An additional horizontal red line was added at $y = 0$ as a helping line. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others. If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason. No subject was shown the same lineup twice. Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected. In every block of 20 lineups that presented to a subject, two lineups with obvious visual patterns were included as attention checks. A subject's submission was only accepted if the actual data plot was identified for at least one attention check. Data of rejected submissions were discarded automatically to maintain the overall data quality.

### 2022-09-28 paper_comparison/paper_comparison/paper_comparison.tex
modify exp design

As pointed out by \citet{vanderplas2021statistical}, the binomial model
doesn't take into account the possible dependencies in the visual test
due to repeated evaluations of the same lineup. And it is inapplicable
to visual test where subjects are asked to select one or more ``most
different'' plots from the lineup. They summarized three common
scenarios in visual inference: (1) \(K\) different lineups are shown to
\(K\) subjects, (2) \(K\) lineups with different null plots but the same
actual data plot are shown to \(K\) subjects, and (3) the same lineup is
shown to \(K\) subjects. Out of these three scenarios, Scenario 3 is the
most common in previous studies as it puts the least constraints on the
experimental design. For Scenario 3, \citet{vanderplas2021statistical}
modelled the probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, they decided to add \(1/s_j\) to
\(c_i\) instead of one, where \(s_j\) is the number of plots subject
\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Thus,
the visual p-value followed by the above model is given as

@@

Instead, we modelled the probability of a lineup being rejected directly
using a logistic regression with the natural logarithm of the effect
size as the only regressor. The same model was applied to different
subsets of the survey data to make comparison of the power.

@@

\hypertarget{motivation-and-overview}{%
\subsection{Motivation and overview}\label{motivation-and-overview}}

With our knowledge, the effectiveness of visual test relative to the
equivalent conventional test in regression diagnostics has not been
validated. Particularly, its ability to detect non-linearity and
heteroskedasticity compared to model structural test and Breusch--Pagan
test. Meanwhile, as mentioned in Section \ref{introduction} section, we
keen to understand the differences between conventional hypothesis
testing and graphical diagnostics in the application of linear
regression diagnostics and why conventional tests are not preferred by
data analysts in linear regression diagnostics. Attempting to find the
answer to these questions, three experiments were conducted.

The experiment I has ideal scenario for conventional testing, where the
visual test is not expected to outperform the conventional test. The
experiment II is a scenario where the conventional test is an
approximate test, in which the visual test may have a chance to match
the performance of the conventional test. The experiment III is designed
for collecting human responses to lineup with only good residual plots
such that the parameter \(\alpha\) in Equation
\ref{eq:pvalue-beta-binomial} can be estimated.

Subjects for all three experiments were recruited from an crowdsourcing
platform called Prolific {[}ref here{]}. Prescreening procedure was
applied during the recruitment, subjects were required to be fluent in
English, with \(98\%\) minimum approval rate in other studies and 10
minimum submissions.

During the experiment, every subject was presented with a block of 20
lineups. For each lineup, the actual data plot was drawn as a standard
residual plot of the null model with raw residuals on the y-axis and
fitted values on the x-axis. An additional horizontal red line was added
at \(y = 0\) as a helping line. The 19 null datasets were generated by
the residual rotation technique, and plotted in the same way. The lineup
consisted of 20 residual plots with one randomly placed actual data
plot. And for every lineup, the subject was asked to select one or more
plots that are most different from others, provide a reason for their
selections, and evaluate how different they think the selected plots
were from others. If there was no noticeable difference between plots in
a lineup, subjects were permitted to select zero plots without providing
the reason. No subject was shown the same lineup twice. Information
about preferred pronoun, age group, education, and previous experience
in visual experiment were also collected.

In every block of 20 lineups that presented to a subject, two lineups
with obvious visual patterns were included as attention checks. A
subject's submission was only accepted if the actual data plot was
identified for at least one attention check. Data of rejected
submissions were discarded automatically to maintain the overall data
quality.

Overall, we planned to collect 7974 evaluations on 1152 uniqued lineups
performed by 443 subjects.

@@

where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
\(\boldsymbol{z}_{raw}\) are vector of size \(n\), \(He_{j}(.)\) is the
\(j\)th-order probabilist's Hermite polynomials,
\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
the random vector to be \(\{-k, k\}\) defined as

@@

Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
out by the null regression, which will result in model misspecification.

@@

where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\) are vector of size \(n\) and \(g(.)\) is
the scaling function defined in \ref{eq:scaling-function}.

@@

For \(b \neq 0\), the variance-covariance matrix of the error term
\(\boldsymbol{\varepsilon}\) is correlated with the regressor \(x\),
which will lead to the presence of heteroskedasticity.

@@

\begin{quote}
Experiment I
\end{quote}

\begin{itemize}
\tightlist
\item
  Treatments for nonlinearity experiment (plots summarising levels)
\end{itemize}

Data of experiment I were simulated using four different order of
probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three different
sample sizes (n = 50, 100, 300), four different standard deviations of
the error (\(\sigma\) = 0.5, 1, 2, 4) and four different distribution of
\(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
parameters used in this experiment is given in Table
\ref{tab:parameter-table}.

\caption{\label{tab:parameter-table}Parameter values for $n$, $j$ $\sigma$, $X_{raw}$}
\centering
\begin{tabular}[t]{rrrl}
\toprule
\multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} & \multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Error SD\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4}
50 & 2 & 0.5 & $U(-1, 1)$\\
100 & 3 & 1.0 & $N(0, 0.3^2)$\\
300 & 6 & 2.0 & $lognormal(0, 0.6^2)/3$\\
 & 18 & 4.0 & $U\{1, 5\}$\\
\bottomrule
\end{tabular}
\end{table}

The range of \(\sigma\), which is a factor controlling the strength of
the signal, was chosen so that different difficulty levels of lineups
were generated, and therefore, the estimated power curve would be smooth
and continuous. The example residual plots are given in Figure
\ref{fig:different-sigma}.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1} \caption{Example residual plots of the linear regression model used in experiment I. Four different values of $\sigma$ are included in the experiment to control the strength of the signal.}\label{fig:different-sigma}
\end{figure}

Four different distribution were used to generate \(X_{raw}\) shown in
Figure \ref{fig:different-dist}. The uniform and the normal distribution
are symmetric and commonly assumed in statistical models. The adjusted
log-normal distribution provides skewed density. And the discrete
uniform distribution provides discreteness in residual plot, which could
enrich the pool of visual patterns.

Figure \ref{fig:example-lineup} shows one of the lineups used in
experiment I. This lineup was produced under \(n = 300\), \(j = 6\),
\(\sigma = 1\) and \(X_{raw} \sim N(0.0.3^2)\). The actual data plot
location was four. All five subjects correctly identified the actual
data plot for this lineup.

\begin{quote}
Experiment II
\end{quote}

\begin{itemize}
\tightlist
\item
  Treatments for heteroskedasticity experiment (plots summarising
  levels)
\end{itemize}

Experiment data were simulated using three different shapes (\(a\) = -1,
0, 1), five different values of \(b\) (b = 0.25, 1, 4, 16, 64), three
different sample sizes (n = 50, 100, 300) and four different
distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
parameters used in this experiment is given in Table
\ref{tab:parameter-table}.

The values of \(a\) was chosen so that different shapes of
heteroskedasticity were included in the residual plot. These include
triangle shape, butterfly shape and inverse triangle shape.

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}

The range of \(b\), which is a factor controlling the strength of the
signal, was chosen so that different difficulty levels of lineups were
generated, and therefore, the estimated power curve would be smooth and
continuous.

@@

Three replications are made for each of the parameter values shown in
Table \ref{tab:parameter-table} resulting in 192 different lineups. For
each lineup, the actual data plot was drawn as a standard residual plot
of the null model with raw residuals on the y-axis and fitted values on
the x-axis. The 19 null datasets were generated by the residual rotation
technique, and plotted in the same way. The lineup consisted of 20
residual plots with one randomly placed actual data plot. Figure
\ref{fig:example-lineup} is an example of one of these lineups. It was
produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
five subjects correctly identified the actual data plot for this lineup.

In addition, each lineup is designed to be evaluated by five different
subjects to provide reasonable estimates of the visual p-value. Thus,
\(192 \times 3 \times 5 / (20 - 2) = 160\) subjects were recruited to
satisfy the design of the experiment I.

@@

>>>>

As pointed out by \citet{vanderplas2021statistical}, the binomial model
doesn't take into account the possible dependencies in the visual test
due to repeated evaluations of the same lineup. And it is inapplicable
to visual test where subjects are asked to select one or more ``most
different'' plots from the lineup. They summarized three common
scenarios in visual inference: (1) \(K\) different lineups are shown to
\(K\) subjects, (2) \(K\) lineups with different null plots but the same
actual data plot are shown to \(K\) subjects, and (3) the same lineup is
shown to \(K\) subjects. Out of these three scenarios, Scenario 3 is the
most common in previous studies as it puts the least constraints on the
experimental design. For Scenario 3, \citet{vanderplas2021statistical}
modelled the probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
instead of one, where \(s_j\) is the number of plots subject \(j\)
selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Since we are
only interested in the selections of the actual data plot \(i\), the
marginal model can be simplified to a beta-binomial model and thus the
visual p-value is given as

@@

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\item
  Instead, we modelled the probability of a lineup being rejected
  directly using a logistic regression with the natural logarithm of the
  effect size as the only regressor. The same model was applied to
  different subsets of the survey data for making comparison of the
  power.
\item
  Instead, we modelled the plot selection \(c_i\) for the actual data
  plot \(i\).
\end{enumerate}

@@

\hypertarget{overview}{%
\subsection{Overview}\label{overview}}

Three experiments were conducted to investigate the difference between
conventional hypothesis testing and graphical diagnostics in the
application of linear regression diagnostics. The experiment I has ideal
scenario for conventional testing, where the visual test is not expected
to outperform the conventional test. The experiment II is a scenario
where the conventional test is an approximate test, in which the visual
test may have a chance to match the performance of the conventional
test. The experiment III is designed for collecting human responses to
lineup with only good residual plots such that the parameter \(\alpha\)
in Equation \ref{eq:pvalue-beta-binomial} can be estimated. Overall, we
planned to collect 7974 evaluations on 1152 uniqued lineups performed by
443 subjects throughout three experiment.

@@

where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\), \(\boldsymbol{x}_{raw}\),
\(\boldsymbol{z}_{raw}\) are vectors of size \(n\), \(He_{j}(.)\) is the
\(j\)th-order probabilist's Hermite polynomials,
\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\), and
\(g(\boldsymbol{x}, k)\) is a scaling function to enforce the support of
the random vector to be \(\{-k, k\}\) defined as

@@

Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
out by the null regression, which will result in model misspecification.
Visual patterns of non-linearity were simulated using four different
order of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)) and
four different distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2)
\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\).

Four different distribution were used to generate \(X_{raw}\) as shown
in Figure \ref{fig:different-dist}. The uniform and the normal
distribution are symmetric and commonly assumed in statistical models.
The adjusted log-normal distribution provides skewed density. And the
discrete uniform distribution provides discreteness in residual plot,
which could enrich the pool of visual patterns.

Figure \ref{fig:example-lineup} shows one of the lineups used in
experiment I. This lineup was produced under \(j = 6\) and
\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
five subjects correctly identified the actual data plot for this lineup.

@@

where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\) are vectors of size \(n\) and \(g(.)\) is
the scaling function defined in \ref{eq:scaling-function}.

@@

For \(b \neq 0\), the variance-covariance matrix of the error term
\(\boldsymbol{\varepsilon}\) is correlated with the regressor \(x\),
which will lead to the presence of heteroskedasticity. Visual patterns
were simulated using three different shapes (\(a\) = -1, 0, 1) and the
same four different distribution of \(X_{raw}\) used in experiment I.

The values of \(a\) was chosen so that different shapes of
heteroskedasticity were included in the residual plot. These include
left-triangle shape, butterfly shape and right-triangle shape as
displayed in Figure \ref{fig:different-shape-of-heter}.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1} \caption{Example residual plots of the linear regression model used in experiment II. Three different shapes (a = -1, 0, 1) are used in the experiment to create left-triangle shape, butterfly shape and right-triangle shape.}\label{fig:different-shape-of-heter}
\end{figure}

@@

To control the strength of the signal, two parameters were used
including \(\sigma\) and \(b\) for controlling the standard deviation of
the error term in experiment I and II respectively.

@@

@@

Subjects for all three experiments were recruited from an crowdsourcing
platform called Prolific {[}ref here{]}. Prescreening procedure was
applied during the recruitment, subjects were required to be fluent in
English, with \(98\%\) minimum approval rate in other studies and 10
minimum submissions. During the experiment, every subject was presented
with a block of 20 lineups. For each lineup, the actual data plot was
drawn as a standard residual plot of the null model with raw residuals
on the y-axis and fitted values on the x-axis. An additional horizontal
red line was added at \(y = 0\) as a helping line. The 19 null datasets
were generated by the residual rotation technique, and plotted in the
same way. The lineup consisted of 20 residual plots with one randomly
placed actual data plot. And for every lineup, the subject was asked to
select one or more plots that are most different from others, provide a
reason for their selections, and evaluate how different they think the
selected plots were from others. If there was no noticeable difference
between plots in a lineup, subjects were permitted to select zero plots
without providing the reason. No subject was shown the same lineup
twice. Information about preferred pronoun, age group, education, and
previous experience in visual experiment were also collected. In every
block of 20 lineups that presented to a subject, two lineups with
obvious visual patterns were included as attention checks. A subject's
submission was only accepted if the actual data plot was identified for
at least one attention check. Data of rejected submissions were
discarded automatically to maintain the overall data quality.
